Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-08-04 Report/introduction.tex
temp updates

The last mile wireless connectivity poses significant bottleneck in the overall data traffic for the interconnected networks. The main challenges in the wireless networks are due to the scarcity of the available resources either in terms of power or spectrum usage and the complexity of the receiver algorithms which is proportional to the mobile battery drain. In order to over come the receiver complexity, \ac{OFDM} based transmissions are introduced for the wideband transmissions. To improve the data rate, multiple antennas are installed at the \ac{BS} and at the user terminals to avail the additional freedom in the form of spatial dimension. The inclusion of \ac{MIMO} technique in the wireless network provides increased data rate or lower outage for the same transmission power and the channel use.

In a network with multiple \acp{BS} serving multiple users (\acs{MU}), the main driving factor for the transmission are the packets waiting at each \ac{BS} corresponding to the different users present in the network. These available packets are transmitted over the shared wireless resources in an efficient manner by including the system limitations and constraints. In this work, we consider the user allocation over the space-frequency resources provided by the \ac{MU-MIMO} \ac{OFDM} framework in the downlink broadcast transmission to reduce the number of queued packets corresponding to the users at the \acp{BS}. Since the space-frequency resources are shared by multiple users associated with different \acp{BS}, the problem can be considered as a resource allocation to minimize the number of backlogged packets at the \acp{BS}.

In general, the resource allocation problems are solved by assigning a binary variable to each user indicating the presence or the absence on a particular resource. In contrast to that, we use the transmit beamformers, which are the complex vectors, as a decision variable in determining the presence or the absence of a user on a particular resource. The purpose of using the transmit beamformers for the scheduling is two fold. Firstly, it determines the transmission rate on a certain resource and secondly, by making the transmit beamformer to be a zero vector, the corresponding user will not be scheduled on a certain resource. The queue minimizing network optimization objective is used to design the beamformers across the coordinating \acp{BS}, since the transmissions are guided by the available backlogged packets. To achieve the best performance, we propose a joint resource allocation scheme over the space and frequency dimensions among the coordinating \acp{BS} to minimize the time that the packets stay in the queues prior to the transmission, and, hence, to avoid packet drops as an indirect objective.

The queue minimizing precoder designs are nothing but the \ac{WSRM} problem with the additional rate constraint determined by the number of backlogged packets for each user in the system. The topic of \ac{MIMO} \ac{BC} precoder design has been studied extensively with different performance criteria in the literature. Due to the nonconvex nature of the linear \ac{MIMO} \ac{BC} precoder design problem, it is usually solved by the sequence of approximated convex problems. In \cite{sin_algorithm}, the \ac{MIMO} rate expression is linearized around a fixed point in order to solve the \ac{DC} problem as a concave maximization problem in an iterative manner. Similar approach of solving the \ac{WSRM} problem via arithmetic-geometric inequality is proposed in \cite{tran2012fast}, providing better convergence rate compared to the former.

The relation between the achievable capacity and the \ac{MSE} for the received symbol by using the fixed \ac{MMSE} receivers as shown in \cite{viswanath1999optimal,mse_duality} are also used to solve the \ac{WSRM} objective. In \cite{christensen2008weighted,wmmse_shi}, the \ac{WSRM} problem is reformulated via \ac{MSE}, casting the problem as a convex one for fixed \ac{MMSE} receivers. In this way, the original problem is expressed in terms of the \ac{MSE} weight, precoders, and decoders. Then the problem is solved using an alternating optimization method, i.e., finding a set of variables while the remaining others are fixed. Additional rate constraints based on the \ac{QoS} requirements are included in the \ac{WSRM} objective are solved via \ac{MSE} reformulation in \cite{kaleva2013primal}.

The problem of precoder design for the \ac{MIMO} \ac{BC} system are solved either by using the centralized controller or at each \ac{BS} independently by exchanging limited information via backhaul. The distributed approaches are performed using the primal or by the dual decomposition. In primal decomposition, the coupling interference variables are fixed for each \ac{BS} wise subproblems and solved for the optimal precoders. The fixed interference are then updated by using the subgradients as discussed in \cite{pennanen2011decentralized}. The dual approach controls the distributed subproblems by fixing the interference price for each \ac{BS} as detailed in \cite{tolli2011decentralized}.

By adjusting the weights properly, we can use the \ac{WSRM} schemes to achieve other performance measures. For example, if the weight of each user is set to be inversely proportional to his/her data rate, a \ac{WSRM} scheme can maintain fairness among users. Similarly, the \ac{WSRM} schemes can be used to solve the design problem considered in this paper. Specifically, to find a rate vector that minimizes the number of queued packets, we should assign weights based on the current queue size of users. The queue states can be incorporated to traditional weighted sum rate objective $\sum_k w_k R_k$ by replacing the weight $w_k$ with the corresponding queue state $Q_k$ or a function depending on it \cite{tassiulas,neely2010stochastic}, which is the outcome of minimizing the Lyapunov drift between the current and the future queue states. Similar algorithms are discussed in the networking literature as the backpreassure algorithm, where the differential queues between the source and the destination nodes are used as the weights scaling the transmission rate \cite{georgiadis2006resource}.

Earlier work on the queue minimization problem was addressed in the survey paper \cite{berry2004cross} and in particular \cite{qps_cioffi}, where the problem of optimal power allocation is considered to minimize the number of backlogged packets via \acl{GP} formulation. Since the problem considered in \cite{qps_cioffi} assumes multi-user \ac{OFDM} model without multi-antenna support, the queue minimizing scheduling reduces to the problem of optimal power allocation. In the context of the wireless networks, the backpreassure algorithm has been extended by formulating the wireless transmission rate as a \ac{MIMO} \ac{BC} optimization problem by using the corresponding user queues as the weight for the associated rate \cite{weeraddana2011resource}.

In the traditional \ac{WSRM} formulation, the precoders are designed for the sub-channels independently due to the interference-free \ac{OFDM} transmission. In this work, since the objective is to identify the transmission rates for all the users across the space-frequency resources, the precoders are to be designed jointly across the space-frequency dimension. In order minimize the number of queued packets at the \acp{BS}, the precoders are designed by the proposed \ac{JSFRA} formulation with two different approaches in a centralized manner. We also propose distributed approaches for the centralized problem for designing the precoders independently across the \acp{BS} with the limited information exchange.

The outline of this paper is as follows. In Section \ref{sec-2-3.2}, we propose the system model and the problem formulation for the current queue minimizing precoder design. Existing and the proposed precoder designs for the queue minimizing objective are discussed elaborately in Section \ref{sec-3}. The distributed solutions are outlined in Section \ref{sec-4} followed by the simulation results for both centralized and the distributed algorithms are provided in Section \ref{sec-5}. Conclusions are drawn in Section \ref{sec-6}.

>>>>

The last mile wireless connectivity poses significant bottleneck in the overall data traffic for the interconnected networks. The main challenges in the wireless networks are due to the scarcity of the available resources either in terms of power or spectrum usage and the complexity of the receiver algorithms which are proportional to the mobile battery drain. In order to overcome the receiver complexity, \ac{OFDM} based transmissions are introduced for the wideband transmissions. To improve data rate, multiple antennas are installed at \acp{BS} and/or at user terminals to avail additional freedom in the form of spatial dimension. The inclusion of \ac{MIMO} technique in wireless networks provides higher data rate or lower outage for the same transmission power and bandwidth.

In a network with multiple \acp{BS} serving multiple users (\acs{MU}), the main driving factor for the transmission are the packets waiting at each \ac{BS} corresponding to the different users present in the network. These available packets are transmitted over the shared wireless resources subject to certain system limitations and constraints. In this work, we consider the problem of transmit design over the space-frequency resources provided by a \ac{MU-MIMO} \ac{OFDM} framework in the downlink broadcast transmission to minimize the number of queued packets. Since the space-frequency resources are shared by multiple users associated with different \acp{BS}, the problem of interest can be viewed as a resource allocation to minimize the total number of backlogged packets of all \acp{BS}.

In general, the resource allocation problems are solved by assigning a binary variable to each user indicating the presence or the absence on a particular resource. In contrast to that, we use the transmit beamformers, which are the complex vectors, as a decision variable in determining the presence or the absence of a user on a particular resource. The purpose of using the transmit beamformers for the scheduling is two fold. Firstly, it determines the transmission rate on a certain resource and secondly, by making the transmit beamformer to be a zero vector, the corresponding user will not be scheduled on a certain resource.

In order to reduce the complexity involved in the precoder design problem, linear precoding are assumed at the \acp{BS} and linear detection is used at users. The queue minimizing network optimization objective is used to design the beamformers across the coordinating \acp{BS}, since the transmissions are guided by the available backlogged packets. To achieve the best performance, we propose a joint resource allocation scheme over the space and frequency dimensions among the coordinating \acp{BS} to minimize the time that the packets stay in the queues prior to the transmission, and, hence, to avoid packet drops as an indirect objective.

The queue minimizing precoder designs are closely related to the \ac{WSRM} problem with additional rate constraints determined by the number of backlogged packets for each user in the system (see Section \ref{sec-3} for further discussions). The topics on \ac{MIMO} \ac{BC} precoder design have been studied extensively with different performance criteria in the literature. Due to the nonconvex nature of most, if not all, linear \ac{MIMO} \ac{BC} precoder design problems, the \ac{SCA} method has gradually become a powerful tool to deal with these problems. For example, in \cite{sin_algorithm}, the nonconvex part of the objective is linearized around an operating point in order to solve the \ac{WSRM} problem in an iterative manner. Similar approach of solving the \ac{WSRM} problem via arithmetic-geometric inequality is proposed in \cite{tran2012fast}.

The connection between the achievable capacity and the \ac{MSE} for the received symbol by using the fixed \ac{MMSE} receivers as shown in \cite{viswanath1999optimal,mse_duality} is also used to solve the \ac{WSRM} problem. In \cite{christensen2008weighted,wmmse_shi}, the \ac{WSRM} problem is reformulated via \ac{MSE}, casting the problem as a convex one for fixed \ac{MMSE} receivers. In this way, the original problem is expressed in terms of the \ac{MSE} weight, precoders, and decoders. Then the problem is solved using an alternating optimization method, i.e., finding a set of variables while the remaining others are fixed. Additional rate constraints based on the \ac{QoS} requirements are included in the \ac{WSRM} problem are solved via \ac{MSE} reformulation in \cite{kaleva2013primal}.

The problem of precoder design for the \ac{MIMO} \ac{BC} system are solved either by using a centralized controller or by using decentralized algorithms where each BS handles their own problems and exchange limited information via backhaul. The distributed approaches are based on the primal or dual decomposition \cite{palomar2006tutorial,boyd2011distributed}. In primal decomposition, the so-called coupling interference variables are fixed for the subproblem at each \ac{BS} to find the optimal precoders. The fixed interference are then updated by using the subgradient method as discussed in \cite{pennanen2011decentralized}. The dual approach controls the distributed subproblems by fixing the \emph{`interference price'} for each \ac{BS} as detailed in \cite{tolli2011decentralized}.

By adjusting the weights properly, we can find arbitrary rate-tuple in the rate region of the system that maximizes other performance measures by solving the resulting WSRM problem. For example, if the weight of each user is set to be inversely proportional to his/her data rate, the corresponding problem guarantees fairness among users. As an approximate method, we may assign weights based on the current queue size of users. More specifically, the queue states can be incorporated to traditional weighted sum rate objective $\sum_k w_k R_k$ by replacing the weight $w_k$ with the corresponding queue state $Q_k$ or a function of it,  which is the outcome of minimizing the Lyapunov drift between the current and the future queue states \cite{tassiulas,neely2010stochastic}. In the backpreassure algorithm, the differential queues between the source and the destination nodes are used as the weights scaling the transmission rate \cite{georgiadis2006resource}.

Earlier studies on the queue minimization problem was summarized in the survey paper \cite{berry2004cross}. In particular, the problem of power allocation to minimize the number of backlogged packets was considered in \cite{qps_cioffi} by geometric programming formulations. Since the problem considered in \cite{qps_cioffi} assumed single antenna transmitters and receivers, the queue minimizing problem reduces to the one of optimal power allocation. In the context of wireless networks, the backpreassure algorithm mentioned above was extended in \cite{weeraddana2011resource} by formulating the corresponding user queues as the weights in the \ac{WSRM} problem.

In this paper, we consider the problem of precoder design across the space-frequency resources to minimize the total number of queued packets of all \acp{BS}. For this highly nonconvex problem, we first propose two centralized methods. In the first method, we relax the nonconvex constraint by the first order taylor approximation around an operating point, which is updated in an iterative manner until convergence or to a certain accuracy. In the second method, we reformulate the \ac{JSFRA} problem using the \ac{MSE} equivalence with the rate expression to solve for the optimal precoders. For distributed implementation, we further proposed decentralized approaches based on primal and \ac{ADMM} scheme to identify the precoders independently across the \acp{BS} by exchanging limited information via backhaul. We also proposed an iterative algorithm by solving the \ac{KKT} equations, which can be implemented easily in a distributed manner.

The remainder of this paper is as follows. In Section \ref{sec-2-3.2}, we introduce the system model and the problem formulation for the queue minimizing precoder design. Existing and the proposed precoder designs for the \ac{JSFRA} problem are presented in Section \ref{sec-3}. The distributed solutions are provided in Section \ref{sec-4} followed by the simulation results in Section \ref{sec-5}. Conclusions are drawn in Section \ref{sec-6}.

### 2014-08-04 Report/joint_wsrm_formulation.tex
temp updates

The queue minimizing algorithms are discussed extensively in the networking literature for providing a congestion-free routing between any two nodes\footnote{routers or user terminals} in the network. One such algorithm used extensively is the \emph{backpreassure algorithm}, discussed in detail in \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. The algorithm determines an optimal control policy in the form of rate or resource allocation for the nodes in the network by considering the differential backlogged packets between the source and the destination nodes.
Even though the algorithm is primarily designed for the wired infrastructure, it can be extended to the wireless networks by designing the user rate variable \me{t_k} in accordance to the wireless network.

The \ac{Q-WSRM} formulation extends the \emph{backpreassure algorithm} to the \ac{MIMO} \ac{OFDM} framework, in which the multiple \acp{BS} acts as the source nodes and the user terminals as the receiver nodes. The control policy in the form of transmit precoders are designed with the objective of minimizing the number of queued packets waiting at the \acp{BS}. The weights used in the precoder design algorithm should be a function of the number of backlogged packets corresponding to each user. In order to find the weights, we adopt the Lyapunov drift minimization as discussed in \cite{neely2010stochastic}, where the Lyapunov function is given by the squared sum of the number backlogged packets as
\begin{equation}
\mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \, \sum_{k \in \mc{U}} Q_k(i),
\end{equation}
where \me{\mbf{Q}(i)} denotes the stacked user queues at the \me{\ith{i}} slot. The Lyapunov function provides a measure of congestion in the system, as discussed in \cite[Ch. 3]{neely2010stochastic}. Now the Lyapunov function drift is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} &=& \frac{1}{2} \left [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \right ] \IEEEyessubnumber \label{eqn-3.1.0} \\
&\leq& \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) }, \IEEEyessubnumber \label{drift-exp}
\end{IEEEeqnarray}
where the inequality is due to the upper bound
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t).
\end{equation}
In order to minimize the number of backlogged packets at each instant, minimization of the Lyapunov drift \eqref{eqn-3.1} is carried over all possible control decisions in the form of transmission rates \me{t_k} to users in the system. The Lyapunov drift conditioned on the current backlogged packets \me{\mbfa{Q}(i)} is given by
\begin{IEEEeqnarray}{rCl}
\underset{\mbf{t}}{\text{minimize}} && \Delta(\mbf{Q}(i)) \triangleq \mathbb{E} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
 &\leq& \mathbb{E} \set {\sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) + \mathbb{E}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)}, \IEEEyessubnumber
\end{IEEEeqnarray}
where the second term on the r.h.s is due to i.i.d assumption on the arrivals. We can bound the first term on the r.h.s by a constant \me{B} for all possible control actions taken for a given channel condition \cite{neely2010stochastic}, the resulting problem can be given as
\begin{equation}
\underset{t_k}{\text{maximize}} \quad \sum_{k \in \mc{U}} \, Q_k(i) \, t_k(i),
\label{backpreassure}
\end{equation}
where the expectation operator is dropped since it is maximized by maximizing the function inside in an opportunistic manner.

Now the problem in \eqref{backpreassure} looks similar to the \ac{WSRM} problem if the weights are replaced by the number of backlogged packets corresponding to the users. The above discussed approach is extended for the wireless networks in \cite{weeraddana2011resource}, where the queue weighted sum rate maximization is considered as the objective function to determine the transmit precoders. In order to avoid the excessive  allocation of the resources, we include an additional rate constraint \me{t_k \leq Q_k} to address \me{[x]^+} operation in \eqref{eqn-2a}. With this, the \ac{Q-WSRM} problem for a wireless system is given by
\begin{IEEEeqnarray}{rCl} \label{q_gen_sum}
\underset{\substack{\mvec{M}{k,n},\gamma_{l,k,n} \\ t_{l,k,n},\mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, Q_k \left ( \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n} \right ) \IEEEyessubnumber \label{eqn-3.1.1} \\
\text{subject to.} & \quad & t_{l,k,n} \leq \log_2(1+\gamma_{l,k,n}) \IEEEyessubnumber \label{eqn-3.1.2} \\
& \quad &\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{eqn-3.1.3} \\
& \quad & \sum_{n=1}^N \sum_{l = 1}^L t_{l,k,n} \leq Q_k, \fall k \in \mc{U} \IEEEyessubnumber \label{eqn-3.1.4}
\end{IEEEeqnarray}
where the precoders are associated with the \me{\gamma_{l,k,n}} defined in \eqref{eq:SINR}. By using the number of queued packets as the weights, the resources can be allocated to the user with the more number of backlogged packets, which essentially does the allocation in a greedy manner. As a special case of the problem defined in \eqref{q_gen_sum}, we can formulate the sum rate maximization problem by setting the weights in \eqref{eqn-3.1.1} as unity, leading to the problem as
\begin{IEEEeqnarray}{rCl}\label{gen_sum}
\underset{\substack{\mvec{M}{k,n},\gamma_{l,k,n} \\ t_{l,k,n},\mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n} \IEEEyessubnumber \\
\text{subject to.} & \quad & \eqref{eqn-3.1.1}, \eqref{eqn-3.1.2}, \eqref{eqn-3.1.3} \; \text{and} \; \eqref{eqn-3.1.4}. \IEEEyessubnumber
\end{IEEEeqnarray}

The problem defined in \eqref{gen_sum} provides an efficient queue minimizing approach as compared to \eqref{q_gen_sum}, since the resource allocation is driven by the channel conditions as compared to the number of queued packets as in \eqref{q_gen_sum}. It is to be noted that in both formulations, the resources allocated to the users are limited by the backlogged packets \eqref{eqn-3.1.4}, thereby handling the problem of over allocation explicitly with an additional rate constraint.

>>>>

The queue minimizing algorithms are discussed extensively in the networking literature to provide congestion-free routing between any two nodes in the network.\footnote{routers or user terminals} One such algorithm  is the \emph{backpreassure algorithm}, discussed in detail in \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. The algorithm determines an optimal control policy in the form of rate or resource allocation for the nodes in the network by considering the differential backlogged packets between the source and the destination nodes. Even though the algorithm is primarily designed for the wired infrastructure, it can be extended to the wireless networks by designing the user rate variable \me{t_k} in accordance to the wireless network.

The \ac{Q-WSRM} formulation extends the \emph{backpreassure algorithm} to the \ac{MIMO} \ac{OFDM} framework, in which the multiple \acp{BS} acts as the source nodes and the user terminals as the receiver nodes. The control policy in the form of transmit precoders are designed with the objective of minimizing the number of queued packets waiting at the \acp{BS}. The weights used in the precoder design algorithm should be a function of the number of backlogged packets corresponding to each user. In order to find the weights, we adopt the Lyapunov drift minimization as discussed in \cite{neely2010stochastic}, where the Lyapunov function is given by the squared sum of the number backlogged packets as
\begin{equation}
\mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \, \sum_{k \in \mc{U}} Q_k(i),
\end{equation}
where \me{\mbf{Q}(i)} denotes the stacked user queues at the \me{\ith{i}} slot. The Lyapunov function provides a measure of congestion in the system, as discussed in \cite[Ch. 3]{neely2010stochastic}. Now the Lyapunov function drift is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} &=& \frac{1}{2} \left [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \right ] \IEEEyessubnumber \label{eqn-3.1.0} \\
&\leq& \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) }, \IEEEyessubnumber \label{drift-exp}
\end{IEEEeqnarray}
where the inequality is due to the upper bound
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t).
\end{equation}
In order to minimize the number of backlogged packets at each instant, minimization of the Lyapunov drift \eqref{eqn-3.1} is carried over all possible control decisions in the form of transmission rates \me{t_k} to users in the system. The Lyapunov drift conditioned on the current backlogged packets \me{\mbfa{Q}(i)} is given by
\begin{IEEEeqnarray}{rCl} \label{backpreassure-approx}
\underset{\mbf{t}}{\text{minimize}} && \Delta(\mbf{Q}(i)) \triangleq \mathbb{E} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
 &\leq& \mathbb{E} \set {\sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) + \mathbb{E}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)}, \IEEEyessubnumber
\end{IEEEeqnarray}
where the second term on the r.h.s is due to i.i.d assumption on the arrivals. We can bound the first term on the r.h.s by a constant \me{B} for all possible control actions taken for a given channel condition \cite{neely2010stochastic}.

Now the expression in \eqref{backpreassure-approx} looks similar to the \ac{WSRM} problem if the weights are replaced by the number of backlogged packets corresponding to the users. The above discussed approach is extended for the wireless networks in \cite{weeraddana2011resource}, where the queue weighted sum rate maximization is considered as the objective function to determine the transmit precoders. In order to avoid the excessive  allocation of the resources, we include an additional rate constraint \me{t_k \leq Q_k} to address \me{[x]^+} operation in \eqref{eqn-2a}. With this, the \ac{Q-WSRM} problem for a wireless system is given by
\begin{IEEEeqnarray}{rCl} \label{q_gen_sum}
\underset{\substack{\mvec{M}{k,n}, \mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, Q_k \left ( \sum_{n = 1}^N \sum_{l=1}^L \log_2(1 + \gamma_{l,k,n}) \right ) \IEEEyessubnumber \label{eqn-3.1.1} \\
\text{subject to.} & \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{eqn-3.1.3} \\
& \quad & \sum_{n=1}^N \sum_{l = 1}^L \log_2(1 + \gamma_{l,k,n}) \leq Q_k, \fall k \in \mc{U} \IEEEyessubnumber \label{eqn-3.1.4}
\end{IEEEeqnarray}
where the precoders are associated with the \me{\gamma_{l,k,n}} defined in \eqref{eq:SINR}. By using the number of queued packets as the weights, the resources can be allocated to the user with the more number of backlogged packets, which essentially does the allocation in a greedy manner.

As a special case of the problem defined in \eqref{q_gen_sum}, we can formulate the sum rate maximization problem by setting the weights in \eqref{eqn-3.1.1} as unity, leading to the problem as
\begin{IEEEeqnarray}{rCl}\label{gen_sum}
\underset{\substack{\mvec{M}{k,n},\mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, \sum_{n = 1}^N \sum_{l=1}^L \log_2(1 + \gamma_{l,k,n}) \IEEEyessubnumber \\
\text{subject to.} & \quad & \eqref{eqn-3.1.1}, \eqref{eqn-3.1.3} \; \text{and} \; \eqref{eqn-3.1.4}. \IEEEyessubnumber
\end{IEEEeqnarray}
The problem defined in \eqref{gen_sum} provides a greedy queue minimizing approach as compared to \eqref{q_gen_sum}, since the resource allocation is driven by the channel conditions in comparison with the number of queued packets as in \eqref{q_gen_sum}. Note that in both formulations, the resources allocated to the users are limited by the backlogged packets with an explicit rate constraint defined by \eqref{eqn-3.1.4}.

### 2014-08-04 Report/problem_formulation.tex
temp updates

The problem defined in \eqref{q_gen_sum} ignores the second order term arising from the Lyapunov drift minimization objective by the limiting it to a constant value. Eq. \eqref{eqn-4.2} provides similar expression when the exponent is set to be \me{\ell_{q=2}} as
\begin{equation}\label{pf-1}
\underset{t_k}{\text{minimize}} \, \sum_k \, v_k^2 = \underset{t_k}{\text{minimize}} \, \sum_k \, Q_k^2 - 2 \, Q_k t_k + t_k^2.
\end{equation}
It is evident that \eqref{pf-1} is equivalent to \eqref{backpreassure} if the second order terms are ignored. Limiting \me{t_k^2} by a constant value, the \ac{Q-WSRM} formulation requires the explicit rate constraint \eqref{eqn-3.1.4} to avoid the resource wastage in the form of over allocation. In the current queue deviation formulation, the explicit rate constraint is not needed, since it is handled by the objective function itself. In contrast to the \ac{WSRM} formulation, the \ac{JSFRA} and the \ac{Q-WSRM} problems include the sub-channels jointly to achive an efficient allocation by identifying the optimal space-frequency resource for each user in the system. The queue deviation objective provides an alternate and a more efficient way to perform the resource allocation as compared to the \ac{Q-WSRM} approach.

Before proceeding further, we note that the constraint in \eqref{eqn-4.2} is handled implicitly by the definition of the \me{\ell_q} in the objective of \eqref{eqn-3}. As a proof, suppose that \me{t_k>Q_k} for a certain \me{k} at optimum, i.e., \me{-{v}_k=t_k-Q_k>0}. Then there exists \me{\delta_k>0} such that \me{-{v}^{\prime}_k=t^{\prime}_k-Q_k<-{v}_k} where \me{t^{\prime}_k=t_k-\delta_k}. Since \me{\|\tilde{\mbf{v}}\|_q=\| |\tilde{\mbf{v}}| \|_q=\||-\tilde{\mbf{v}}|\|_q}, this means that the newly created vector \me{\mbf{t}^{\prime}} achieves a smaller objective which contradicts with the fact that an optimal solution has been obtained. The choice of the norm \me{\ell_q} used in the objective function \cite{berry2004cross,qps_cioffi} alters the priorities for the queue deviation function as
\begin{itemize}
\item With \me{\ell_{q = 1}}, the objective results in greedy allocation \textit{i.e}, emptying the queue of users with good channel condition before considering the users with worse channel conditions. As a special case, it is easy to see that \eqref{eqn-3} reduces to the \ac{WSRM} problem \eqref{gen_sum} when the queue size is large enough for all users.
\item With \me{\ell_{q = 2}}, the objective prioritizes users with higher number of queued packets before considering the users with a smaller number of backlogged packets. For example, it could be more ideal for the delay limited scenario when the packet arrival rates of the users are similar, since the backlogged packets is proportional to the delay in the transmission following the Little's law \cite{neely2010stochastic}.
\item With \(\ell_{q = \infty} \), the objective minimizes the maximum number of queued packets among users with the current transmission, thereby providing queue fairness by allocating the resources proportional to the number of backlogged packets.
\end{itemize}

>>>>

The problem defined in \eqref{q_gen_sum} ignores the second order term arising from the Lyapunov drift minimization objective by the limiting it to a constant value. Eq. \eqref{eqn-4.2} provides similar expression when the exponent is set to be \me{\ell_{q=2}} as
\begin{equation}\label{pf-1}
\underset{t_k}{\text{minimize}} \, \sum_k \, v_k^2 = \underset{t_k}{\text{minimize}} \, \sum_k \, Q_k^2 - 2 \, Q_k t_k + t_k^2.
\end{equation}
It is evident that \eqref{pf-1} is equivalent to \eqref{backpreassure-approx} if the second order terms are ignored. Limiting \me{t_k^2} by a constant value, the \ac{Q-WSRM} formulation requires the explicit rate constraint \eqref{eqn-3.1.4} to avoid the resource wastage in the form of over allocation. In the current queue deviation formulation, the explicit rate constraint is not needed, since it is handled by the objective function itself. In contrast to the \ac{WSRM} formulation, the \ac{JSFRA} and the \ac{Q-WSRM} problems include the sub-channels jointly to achive an efficient allocation by identifying the optimal space-frequency resource for each user in the system. The queue deviation objective provides an alternate and a more efficient way to perform the resource allocation as compared to the \ac{Q-WSRM} approach.
